Long groovy title goes here

### 2013-06-25 shanghai.tex
\EEx

@@

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' in this
paper. Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$ \David{I think our notation is clearer if we use $k$
  rather than $\Fq$}; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : G\times G \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob_{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob_{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

@@

The main result of this paper is a canonical isomorphism
\[
\CSiso{\textswab{T}_d} 
\iso
\Hom_{< d}(T(K),\EE^\times)
\] 
between the Picard group of the category of character sheaves on
$\textswab{T}_d$ and the group of characters $T(K) \to \EE^\times$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \Clifton{I'll come back to working on
  this when we're closer to being finished. I'd be just as happy if
  you took this cup from my lips, though.}  In particular, by
specializing to the case $T= \Gm{K}$ and by
choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
be), we arrive at a model for the group of quasi-characters of
$K^\times$: the monoidal category of character sheaves on the
Greenberg transform of the Neron model of
$\Gm{K}$.

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob_G$ and
therefore an automorphism $F_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EE^\times$, canonically (this
follows from second condition appearing in the definition of character
sheaves). The isomorphism $\mu : \bm^* \gcs{L} \to \gcs{L}\boxtimes\gcs{L}$
guarantees that the function $t^\Frob_\cs{L} : G(\Fq)\to \EE^\times$,
defined by $g \mapsto F_{\bg}$, is a group homomorphism;
we refer to $t^\Frob_\cs{L}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
\Clifton{There is no actual trace involved, since $\gcs{L}$ is a rank
  $1$ local system. More generally, trace is needed.}

It is clear that $t^\Frob_{\cs{L}\otimes \cs{L'}} = t^\Frob_\cs{L} \cdot t^\Frob_\cs{L'}$
(pointwise multiplication of functions) and that $t^\Frob_\cs{L} = t^\Frob_\cs{L'}$
if $\cs{L} \iso \cs{L'}$ in $\CS(G)$. In this way we obtain a group homomorphism
\[
\cdef{t_{G}} : \CSiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EE^\times),
\qquad \text{defined by}\qquad \cs{L} \mapsto t^\Frob_\cs{L},
\]
from the Picard group of character sheaves on $G$ to the character group of $G(\Fq)$.

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \xymatrix{
  \CSiso{G} \ar[r]^{f^*} \ar[d]^{t_{G}} &  \ar[d]^{t_{H}} \CSiso{H}\\
  \Hom_\text{grp}(G(\Fq),\EE^\times) \ar[r] & \Hom_\text{grp}(H(\Fq),\EE^\times)
  }
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

\begin{proposition}\label{prop:connected-1}
  If $G$ is a geometrically connected commutative algebraic group over $\Fq$ then
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EE^\times),
  \]
  is an isomorphism of groups.
\end{proposition}

@@

  Let $g$ be any $\Fq$-rational point on $G$ (such as the identity for
  $G$) and let $\bg$ be the geometric point on $G$ lying above
  $g$ (we fixed $\Fq \hookrightarrow \bFq$ long ago). Since $G$ is
  geometrically connected, the geometric point $\bg$ determines
  an equivalence between the category of Weil local systems on $G$ and
  $\ell$-adic representations of $\Weil(G,\bg)$
  \cite{Deligne:Weil2}*{1.1.12}. Now let $(\gcs{L},\mu,F)$ be a
  character sheaf on $G$. Then the isomorphism class of the Weil local
  system $(\gcs{L},F)$ corresponds to an $\ell$-adic character of the
  Weil group $\Weil(G, \bg)$. The $\Fq$-rational point $g$ under
  the geometric point $\bg$ determines a splitting
  $\Weil(\bFq/\Fq)\to \Weil(G,\bg)$ of the exact sequence
  \[
  \xymatrix{
    1 \ar[r] & \pi_1(\bG, \bg) \ar[r] &  \Weil(G,\bg) \ar[r]
    & \ar@/_1pc/@{.>}[l]_{g} \Weil(\bFq/\Fq) \ar[r] & 1
    }
  \]
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character
  \[
  t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EE^\times \qquad \text{defined by}\qquad \Frob \mapsto t^\Frob_{(\gcs{L},F)}(g)
  \]
  of the Weil group $\Weil(\bFq/\Fq)$; the image of $\Frob$ under this
  character is $t^\Frob_\cs{L}(g)$, the trace of Frobenius of $\cs{L}$ at $g$.

  On the other hand, we have already seen that $t^\Frob_\cs{L} : G(\Fq) \to \EE^\times$
  is a group homomorphism. Since $G$ is an algebraic group over $\Fq$, it is a
  variety over $\Fq$, so finitely generated over $\Fq$, so $G(\Fq)$ is finite;
  it follows that $t^\Frob_\cs{L}(g) = t^\Frob_{(\gcs{L},F)}(g)$ is a root of unity,
  for every $g\in G(\Fq)$.  Since $\Weil(\bFq/\Fq)$ is generated by
  $\Frob$ and $t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EE^\times$ is
  a character, it follows that the image of $t_{(\gcs{L},F)}(g)$ is a
  finite subset of $\mu(\EE)$. Thus, $t_{(\gcs{L},F)}(g)$ extends to
  an $\ell$-adic character of $\Gal(\bFq/\Fq)$, which will be denoted
  by the same symbol.
  \[
  \xymatrix{
    1 \ar[r] & \ar@{=}[d] \pi_1(\bG, \bg) \ar[r] & \ar[d] \Weil(G,\bg) \ar[r] & \ar[d]  \Weil(\bFq/\Fq) \ar[r] & 1\\
    1 \ar[r] &  \pi_1(\bG, \bg) \ar[r] & \pi_1(G,\bg) \ar[r] & \Gal(\bFq/\Fq) \ar[r] & 1\\
    }
  \]

  We may now lift the $\ell$-adic character $t_{(\gcs{L},F)}(g) : \Gal(\bFq/\Fq) \to \EE^\times$
  to an $\ell$-adic character of $\pi_1(G,\bg)$ using the canonical topological group map
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$, above. But $\bg$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and local systems on $G$. Let
  $\cs{L}_0$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b^*\cs{L}_0 \iso \gcs{L}$, so $b^*$ is essentially surjective.

@@

  To finish the proof observe that the functions--sheaves dictionary,
  as it pertains to characters, shows that
  \[
  t_{G} : \CSbiso{G} \to \Hom_\text{grp}(G(\Fq),\EE^\times)
  \]
  is an isomorphism of groups; see \cite{Laumon}*{1.1.3} or \cite{SGA4.5}*{Sommes trigonométriques}.
\end{proof}

@@

\begin{proposition}\label{prop:etale}
  If $G$ is a commutative \'etale group scheme over $\Fq$ and
  $G(\bFq)$ is finitely generated then there is a canonical
  isomorphism
  \[
  \CSiso{G} \longrightarrow \Hh^1(\Weil(\bFq/\Fq), \Hom_\text{grp}(G(\bFq),\EE^\times)),
  \qquad \text{given by}\qquad
  [\cs{L}] \mapsto [t_\cs{L}],
  \]
  where $t_\cs{L}: \Weil(\bFq/\Fq)\to \Hom_\text{grp}(G(\bFq),\EE^\times)$ is a cocycle such that
  \[
  \forall g\in G(\Fq), \qquad t_\cs{L}(\Frob)(g) = t^\Frob_\cs{L}(g).
  \]
\end{proposition}

\begin{proof}
  To simplify notation, we set $W_\Fq \ceq \Weil(\bFq/\Fq)$ and
  $G(\bFq)^* \ceq \Hom_\text{grp}(G(\bFq),\EE^\times)$.  Recall the
  equivalence between etale group schemes over $\Fq$ and the category
  of groups equipped with a continuous action of $\Gal(\bFq/\Fq)$ (for
  the discrete topology on the group), given by $G \mapsto G(\bFq)$.
  Under this equivalence any character sheaf $\cs{L}=(\gcs{L},\mu,F)$
  on $G$ corresponds to:
  \begin{enumerate}
  \labitem{(cs.0)}{cs.0} an indexed set of one-dimensional
    $\EE$-vector spaces $\gcs{L}_x \ceq x^*\gcs{L}$, as $x$ runs over
    $G(\bFq)$;

@@

  We now show that the function $\obj \CS(G(\bFq)) \longrightarrow \Hh^1(W_\Fq,G(\bFq)^*)$
  factors through $\obj \CS(G(\bFq)) \to \CSiso{G(\bFq)}$.
  To see this, let $\varphi : \cs{L} \to \cs{L'}$ be an isomorphism and
  let $t$ and $t'$ be any cocycles representing the classes in the
  image of $\cs{L}$ and $\cs{L'}$ under \eqref{cohomologous}; by construction,
  $t$ and $t'$ are defined by sections $s$ and $s'$. Now, $\varphi_x(s(x))$ is a
  non-zero scalar multiple of $s'(x)$; define $\phi: G(\bFq) \to \EE^\times$
  by $\varphi_x(s(x)) = \phi(x) s'(x)$. By the second part of condition~\ref{cs.3},
  $\phi(x+y) = \phi(x)\phi(y)$, so $\phi \in G(\bFq)^*$. By the first part of
  condition~\ref{cs.3}, $t'(\Frob) = t(\Frob) \cdot (\,^{\Frob^{-1}}\phi \cdot \phi^{-1})$,
  so $t$ and $t'$ are cohomologous. This concludes the definition of
  \begin{equation}\label{pich1}
    \CSiso{G(\bFq} \longrightarrow \Hh^1(W_\Fq,G(\bFq)^*),
    \qquad \text{defined by} \qquad [\cs{L}] \mapsto [t_\cs{L}].
  \end{equation}
  Now it follows from \eqref{t} that the image of the isomorphism
  class of $\cs{L}$ is indeed represented by a cocycle $t_\cs{L}$ such
  that $t_\cs{L}(\Frob) = t^\Frob_\cs{L}$ on $G(\Fq)$, which is to
  say, when evaluated at $\Frob$, the cocycle $t_\cs{L}$ is an
  extension of the trace of Frobenius of $\cs{L}$.

@@

%\input{etale-2}
%Define $\rho \colon \Hh^1(W_{\Fq}, \Hom(G(\bFq), \EE^\times))
% \rightarrow \Hom(G(\Fq), \EE^\times)$ by $\rho(\phi)(g) =
% \phi(\Frob)(g)$.

\begin{proposition}\label{prop:X}
  Let $X$ be a finitely generated abelian group equipped with an
  action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on
  $X$.  Then
  \begin{align*}
    \Hh^1(\Fq, \Hom(X, \EE^\times)) &\xrightarrow{\rho} \Hom(\Hh^0(\Fq, X), \EE^\times) \\
    \rho([z])(x) &= z(\Frob)(x)
  \end{align*}
  is an isomorphism of groups.
\end{proposition}
\begin{proof}
  We first prove that $\rho$ is surjective.  Suppose that
  $f \in \Hom(X^\Frob, \EE^\times).$ Since $\EE^\times$ is divisible
  as an abelian group it is injective as a $\ZZ$-module and thus
  $\Ext^1_{\ZZ}(X/X^\Frob, \EE^\times) = 0$.  Applying the functor
  $\Hom(\mbox{---}, \EE^\times)$ to
  $$0 \rightarrow X^\Frob \rightarrow X \rightarrow X/X^\Frob \rightarrow 0$$
  and considering the associated long exact sequence gives the
  existence of an $\tilde{f} \in \Hom(X, \EE^\times)$ restricting to
  $f$.  Define a $1$-cocycle $z \in \Hh^1(\Fq, \Hom(X,\EE^\times))$ by
  mapping $\Frob$ to $\tilde{f}$.  Then $\rho([z]) = f$.

  Now suppose that $[z] \in \Hh^1(\Fq,\Hom(X,\EE^\times))$ with
  $\rho([z]) = 1$; write $\tilde{f}$ for $z(\Frob)$.  The injectivity
  of $\rho$ is equivalent to the existence of a homomorphism $\psi
  \colon X \rightarrow \EE^\times$ with
  $$\tilde{f}(x) = \frac{\psi(\Frob(x))}{\psi(x)}$$
  for $x \in X$.  In fact, we may replace the construction of $\psi$
  for all $\ZZ[\Gal(\bFq/\Fq)]$-modules $X$ with the construction of
  analogous $\psi$s for a different class of modules.

  Since the action of $\Gal(\bFq/\Fq)$ is continuous, there is an $m$
  so that $\Frob^m$ acts trivially on $X$; write $\Fqm$ for the degree
  $m$ extension of $\Fq$ and $\Gamma_m$ for $\Gal(\Fqm/\Fq)$.  Since
  $T^m-1$ factors as a squarefree product of cyclotomic polynomials,
  the Chinese remainder theorem expresses $\ZZ[\Gamma_m]$ as a product
  of $\ZZ[\zeta_d]$ for $d$ dividing $m$.  Each $\ZZ[\zeta_d]$ is the
  maximal order in a number field and thus $\ZZ[\Gamma_m]$ is a
  product of Dedekind domains.  Therefore $X$ decomposes into a direct
  sum of cyclic $\ZZ[\Gamma_m]$-modules \cite[?]{Brandal}, each of the
  form $\ZZ[\Gamma_m] / I$ for some ideal $I \subset \ZZ[\Gamma_m]$.
  The ideal $I$ corresponds to a direct product of ideals $I_d$ in the
  $\ZZ[\zeta_d]$, and within each $\ZZ[\zeta_d]$ the Chinese remainder
  theorem allows a further decomposition of $\ZZ[\zeta_d] / I_d$ as a
  direct sum of $\ZZ[\zeta_d]$-modules of the form $\ZZ[\zeta_d]/P^s$.
  So if we can construct, for every cyclotomic ring $\ZZ[\zeta_d]$,
  prime ideal $P \subset \ZZ[\zeta_d]$ and positive integer $s$, a
  $\psi \colon \ZZ[\zeta_d]/P^s \rightarrow \EE^\times$ with
  \begin{equation} \label{eq:psi-condition}
    \tilde{f}(x) = \frac{\psi(\zeta_d \cdot x)}{\psi(x)},
  \end{equation}
  then the existence of the original $\psi$ follows because $X$ is a
  direct sum of such $\ZZ[\zeta_d]/P^s$.

  We now fix a $d$, $P$ and $s$, write $\zeta$ for $\zeta_d$ and
  choose a polynomial
  $$h(T) = a_rT^r + a_{r-1}T^{r-1} + \cdots + a_0$$
  so that $P^s = (p^s, h(\zeta_d))$.
  % Setting $\psi = 1$ will satisfy equation \ref{eq:psi-condition}
  % when $d = 1$.
  An an abelian group, the quotient $\ZZ[\zeta] / P^s$ is generated by
  $\zeta^i$ for $0 \le i < \phi(d)$; we define
  \begin{equation} \label{eq:psi-def}
    \psi(\zeta^i) = \alpha \prod_{j=0}^{i-1} \tilde{f}(\zeta^j)
  \end{equation}
  for an $\alpha \in \EE^\times$ to be chosen later.  Since there are
  relations among the $\zeta^i$ we first need to check that $\psi$ is
  well-defined.

>>>>

\newcommand{\EEx}{\EE^\times}

@@

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' in this
paper. Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$ \David{I think our notation is clearer if we use $k$
  rather than $\Fq$}; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob_{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob_{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

@@

The main result of this paper is a canonical isomorphism
\[
\CSiso{\textswab{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\textswab{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \Clifton{I'll come back to working on
  this when we're closer to being finished. I'd be just as happy if
  you took this cup from my lips, though.}  In particular, by
specializing to the case $T= \Gm{K}$ and by
choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
be), we arrive at a model for the group of quasi-characters of
$K^\times$: the monoidal category of character sheaves on the
Greenberg transform of the Neron model of
$\Gm{K}$.

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob_G$ and
therefore an automorphism $F_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically (this
follows from second condition appearing in the definition of character
sheaves). The isomorphism $\mu : \bm^* \gcs{L} \to \gcs{L}\boxtimes\gcs{L}$
guarantees that the function $t^\Frob_\cs{L} : G(\Fq)\to \EEx$,
defined by $g \mapsto F_{\bg}$, is a group homomorphism;
we refer to $t^\Frob_\cs{L}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
\Clifton{There is no actual trace involved, since $\gcs{L}$ is a rank
  $1$ local system. More generally, trace is needed.}

It is clear that $t^\Frob_{\cs{L}\otimes \cs{L'}} = t^\Frob_\cs{L} \cdot t^\Frob_\cs{L'}$
(pointwise multiplication of functions) and that $t^\Frob_\cs{L} = t^\Frob_\cs{L'}$
if $\cs{L} \iso \cs{L'}$ in $\CS(G)$. In this way we obtain a group homomorphism
\[
\cdef{t_{G}} : \CSiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto t^\Frob_\cs{L},
\]
from the Picard group of character sheaves on $G$ to the character group of $G(\Fq)$.

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \xymatrix{
  \CSiso{G} \ar[r]^{f^*} \ar[d]^{t_{G}} &  \ar[d]^{t_{H}} \CSiso{H}\\
  \Hom_\text{grp}(G(\Fq),\EEx) \ar[r] & \Hom_\text{grp}(H(\Fq),\EEx)
  }
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

\begin{proposition}\label{prop:connected-1}
  If $G$ is a geometrically connected commutative algebraic group over $\Fq$ then
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx),
  \]
  is an isomorphism of groups.
\end{proposition}

@@

  Let $g$ be any $\Fq$-rational point on $G$ (such as the identity for
  $G$) and let $\bg$ be the geometric point on $G$ lying above
  $g$ (we fixed $\Fq \hookrightarrow \bFq$ long ago). Since $G$ is
  geometrically connected, the geometric point $\bg$ determines
  an equivalence between the category of Weil local systems on $G$ and
  $\ell$-adic representations of $\Weil(G,\bg)$
  \cite{Deligne:Weil2}*{1.1.12}. Now let $(\gcs{L},\mu,F)$ be a
  character sheaf on $G$. Then the isomorphism class of the Weil local
  system $(\gcs{L},F)$ corresponds to an $\ell$-adic character of the
  Weil group $\Weil(G, \bg)$. The $\Fq$-rational point $g$ under
  the geometric point $\bg$ determines a splitting
  $\Weil(\bFq/\Fq)\to \Weil(G,\bg)$ of the exact sequence
  \[
  \xymatrix{
    1 \ar[r] & \pi_1(\bG, \bg) \ar[r] &  \Weil(G,\bg) \ar[r]
    & \ar@/_1pc/@{.>}[l]_{g} \Weil(\bFq/\Fq) \ar[r] & 1
    }
  \]
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character
  \[
  t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EEx \qquad \text{defined by}\qquad \Frob \mapsto t^\Frob_{(\gcs{L},F)}(g)
  \]
  of the Weil group $\Weil(\bFq/\Fq)$; the image of $\Frob$ under this
  character is $t^\Frob_\cs{L}(g)$, the trace of Frobenius of $\cs{L}$ at $g$.

  On the other hand, we have already seen that $t^\Frob_\cs{L} : G(\Fq) \to \EEx$
  is a group homomorphism. Since $G$ is an algebraic group over $\Fq$, it is a
  variety over $\Fq$, so finitely generated over $\Fq$, so $G(\Fq)$ is finite;
  it follows that $t^\Frob_\cs{L}(g) = t^\Frob_{(\gcs{L},F)}(g)$ is a root of unity,
  for every $g\in G(\Fq)$.  Since $\Weil(\bFq/\Fq)$ is generated by
  $\Frob$ and $t_{(\gcs{L},F)}(g) : \Weil(\bFq/\Fq) \to \EEx$ is
  a character, it follows that the image of $t_{(\gcs{L},F)}(g)$ is a
  finite subset of $\mu(\EE)$. Thus, $t_{(\gcs{L},F)}(g)$ extends to
  an $\ell$-adic character of $\Gal(\bFq/\Fq)$, which will be denoted
  by the same symbol.
  \[
  \xymatrix{
    1 \ar[r] & \ar@{=}[d] \pi_1(\bG, \bg) \ar[r] & \ar[d] \Weil(G,\bg) \ar[r] & \ar[d]  \Weil(\bFq/\Fq) \ar[r] & 1\\
    1 \ar[r] &  \pi_1(\bG, \bg) \ar[r] & \pi_1(G,\bg) \ar[r] & \Gal(\bFq/\Fq) \ar[r] & 1\\
    }
  \]

  We may now lift the $\ell$-adic character $t_{(\gcs{L},F)}(g) : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character of $\pi_1(G,\bg)$ using the canonical topological group map
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$, above. But $\bg$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and local systems on $G$. Let
  $\cs{L}_0$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b^*\cs{L}_0 \iso \gcs{L}$, so $b^*$ is essentially surjective.

@@

  To finish the proof observe that the functions--sheaves dictionary,
  as it pertains to characters, shows that
  \[
  t_{G} : \CSbiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups; see \cite{Laumon}*{1.1.3} or \cite{SGA4.5}*{Sommes trigonométriques}.
\end{proof}

@@

\begin{proposition}\label{prop:etale}
  If $G$ is a commutative \'etale group scheme over $\Fq$ and
  $G(\bFq)$ is finitely generated then there is a canonical
  isomorphism
  \[
  \CSiso{G} \longrightarrow \Hh^1(\Weil(\bFq/\Fq), \Hom_\text{grp}(G(\bFq),\EEx)),
  \qquad \text{given by}\qquad
  [\cs{L}] \mapsto [t_\cs{L}],
  \]
  where $t_\cs{L}: \Weil(\bFq/\Fq)\to \Hom_\text{grp}(G(\bFq),\EEx)$ is a cocycle such that
  \[
  \forall g\in G(\Fq), \qquad t_\cs{L}(\Frob)(g) = t^\Frob_\cs{L}(g).
  \]
\end{proposition}

\begin{proof}
  To simplify notation, we set $W_\Fq \ceq \Weil(\bFq/\Fq)$ and
  $G(\bFq)^* \ceq \Hom_\text{grp}(G(\bFq),\EEx)$.  Recall the
  equivalence between etale group schemes over $\Fq$ and the category
  of groups equipped with a continuous action of $\Gal(\bFq/\Fq)$ (for
  the discrete topology on the group), given by $G \mapsto G(\bFq)$.
  Under this equivalence any character sheaf $\cs{L}=(\gcs{L},\mu,F)$
  on $G$ corresponds to:
  \begin{enumerate}
  \labitem{(cs.0)}{cs.0} an indexed set of one-dimensional
    $\EE$-vector spaces $\gcs{L}_x \ceq x^*\gcs{L}$, as $x$ runs over
    $G(\bFq)$;

@@

  We now show that the function $\obj \CS(G(\bFq)) \longrightarrow \Hh^1(W_\Fq,G(\bFq)^*)$
  factors through $\obj \CS(G(\bFq)) \to \CSiso{G(\bFq)}$.
  To see this, let $\varphi : \cs{L} \to \cs{L'}$ be an isomorphism and
  let $t$ and $t'$ be any cocycles representing the classes in the
  image of $\cs{L}$ and $\cs{L'}$ under \eqref{cohomologous}; by construction,
  $t$ and $t'$ are defined by sections $s$ and $s'$. Now, $\varphi_x(s(x))$ is a
  non-zero scalar multiple of $s'(x)$; define $\phi: G(\bFq) \to \EEx$
  by $\varphi_x(s(x)) = \phi(x) s'(x)$. By the second part of condition~\ref{cs.3},
  $\phi(x+y) = \phi(x)\phi(y)$, so $\phi \in G(\bFq)^*$. By the first part of
  condition~\ref{cs.3}, $t'(\Frob) = t(\Frob) \cdot (\,^{\Frob^{-1}}\phi \cdot \phi^{-1})$,
  so $t$ and $t'$ are cohomologous. This concludes the definition of
  \begin{equation}\label{pich1}
    \CSiso{G(\bFq} \longrightarrow \Hh^1(W_\Fq,G(\bFq)^*),
    \qquad \text{defined by} \qquad [\cs{L}] \mapsto [t_\cs{L}].
  \end{equation}
  Now it follows from \eqref{t} that the image of the isomorphism
  class of $\cs{L}$ is indeed represented by a cocycle $t_\cs{L}$ such
  that $t_\cs{L}(\Frob) = t^\Frob_\cs{L}$ on $G(\Fq)$, which is to
  say, when evaluated at $\Frob$, the cocycle $t_\cs{L}$ is an
  extension of the trace of Frobenius of $\cs{L}$.

@@

%\input{etale-2}
%Define $\rho \colon \Hh^1(W_{\Fq}, \Hom(G(\bFq), \EEx))
% \rightarrow \Hom(G(\Fq), \EEx)$ by $\rho(\phi)(g) =
% \phi(\Frob)(g)$.

\begin{proposition}\label{prop:X}
  Let $X$ be a finitely generated abelian group equipped with an
  action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on
  $X$.  Then
  \begin{align*}
    \Hh^1(\Fq, \Hom(X, \EEx)) &\xrightarrow{\rho} \Hom(\Hh^0(\Fq, X), \EEx) \\
    \rho([z])(x) &= z(\Frob)(x)
  \end{align*}
  is an isomorphism of groups.
\end{proposition}
\begin{proof}
  We first prove that $\rho$ is surjective.  Suppose that
  $f \in \Hom(X^\Frob, \EEx).$ Since $\EEx$ is divisible
  as an abelian group it is injective as a $\ZZ$-module and thus
  $\Ext^1_{\ZZ}(X/X^\Frob, \EEx) = 0$.  Applying the functor
  $\Hom(\mbox{---}, \EEx)$ to
  $$0 \rightarrow X^\Frob \rightarrow X \rightarrow X/X^\Frob \rightarrow 0$$
  and considering the associated long exact sequence gives the
  existence of an $\tilde{f} \in \Hom(X, \EEx)$ restricting to
  $f$.  Define a $1$-cocycle $z \in \Hh^1(\Fq, \Hom(X,\EEx))$ by
  mapping $\Frob$ to $\tilde{f}$.  Then $\rho([z]) = f$.

  Now suppose that $[z] \in \Hh^1(\Fq,\Hom(X,\EEx))$ with
  $\rho([z]) = 1$; write $\tilde{f}$ for $z(\Frob)$.  The injectivity
  of $\rho$ is equivalent to the existence of a homomorphism $\psi
  \colon X \rightarrow \EEx$ with
  $$\tilde{f}(x) = \frac{\psi(\Frob(x))}{\psi(x)}$$
  for $x \in X$.  In fact, we may replace the construction of $\psi$
  for all $\ZZ[\Gal(\bFq/\Fq)]$-modules $X$ with the construction of
  analogous $\psi$s for a different class of modules.

  Since the action of $\Gal(\bFq/\Fq)$ is continuous, there is an $m$
  so that $\Frob^m$ acts trivially on $X$; write $\Fqm$ for the degree
  $m$ extension of $\Fq$ and $\Gamma_m$ for $\Gal(\Fqm/\Fq)$.  Since
  $T^m-1$ factors as a squarefree product of cyclotomic polynomials,
  the Chinese remainder theorem expresses $\ZZ[\Gamma_m]$ as a product
  of $\ZZ[\zeta_d]$ for $d$ dividing $m$.  Each $\ZZ[\zeta_d]$ is the
  maximal order in a number field and thus $\ZZ[\Gamma_m]$ is a
  product of Dedekind domains.  Therefore $X$ decomposes into a direct
  sum of cyclic $\ZZ[\Gamma_m]$-modules \cite[?]{Brandal}, each of the
  form $\ZZ[\Gamma_m] / I$ for some ideal $I \subset \ZZ[\Gamma_m]$.
  The ideal $I$ corresponds to a direct product of ideals $I_d$ in the
  $\ZZ[\zeta_d]$, and within each $\ZZ[\zeta_d]$ the Chinese remainder
  theorem allows a further decomposition of $\ZZ[\zeta_d] / I_d$ as a
  direct sum of $\ZZ[\zeta_d]$-modules of the form $\ZZ[\zeta_d]/P^s$.
  So if we can construct, for every cyclotomic ring $\ZZ[\zeta_d]$,
  prime ideal $P \subset \ZZ[\zeta_d]$ and positive integer $s$, a
  $\psi \colon \ZZ[\zeta_d]/P^s \rightarrow \EEx$ with
  \begin{equation} \label{eq:psi-condition}
    \tilde{f}(x) = \frac{\psi(\zeta_d \cdot x)}{\psi(x)},
  \end{equation}
  then the existence of the original $\psi$ follows because $X$ is a
  direct sum of such $\ZZ[\zeta_d]/P^s$.

  We now fix a $d$, $P$ and $s$, write $\zeta$ for $\zeta_d$ and
  choose a polynomial
  $$h(T) = a_rT^r + a_{r-1}T^{r-1} + \cdots + a_0$$
  so that $P^s = (p^s, h(\zeta_d))$.
  % Setting $\psi = 1$ will satisfy equation \ref{eq:psi-condition}
  % when $d = 1$.
  An an abelian group, the quotient $\ZZ[\zeta] / P^s$ is generated by
  $\zeta^i$ for $0 \le i < \phi(d)$; we define
  \begin{equation} \label{eq:psi-def}
    \psi(\zeta^i) = \alpha \prod_{j=0}^{i-1} \tilde{f}(\zeta^j)
  \end{equation}
  for an $\alpha \in \EEx$ to be chosen later.  Since there are
  relations among the $\zeta^i$ we first need to check that $\psi$ is
  well-defined.
